---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(ImmuneSpaceR)
library(dplyr)
library(tidyr)
library(data.table)
library(plotly)
library(Rtsne)
```

```{r params}
  # Non-User Selected Params
  baseUrl <- labkey.url.params$baseUrl
  sdy <- gsub("/Studies/", "", labkey.url.params$folderPath)
  
  # Main Params
  timeAs <- labkey.url.params$timeAs
  assays <- strsplit(labkey.url.params$assays, ",")[[1]]
  timePts <- strsplit(labkey.url.params$timePts, ",")[[1]]
  plotType <- labkey.url.params$plotType
  
  # additional options
  perplexity <- as.integer(labkey.url.params$perplexity)
  numComponents <- as.integer(labkey.url.params$numComponents)
  impute <- labkey.url.params$impute
  responseVar <- labkey.url.params$responseVar
```

```{r helper-fn}
getAssayData <- function(assay, con, timePts, timeAs){
  if(assay != "gene_expression_files"){
    original_view <- assay != "mbaa"
    dat <- con$getDataset(assay, original_view = original_view, transformMethod = "auto")
  }else{
    eset <- con$getGEMatrix(con$cache$GE_matrices$name)
    dat <- data.frame(t(Biobase::exprs(eset)),stringsAsFactors = F)
    gef <- con$getDataset("gene_expression_files",
                          original_view = TRUE)
    ptid <- gef$participant_id[ match(rownames(dat), gef$biosample_accession) ]
    dat$study_time_collected <- gef$study_time_collected[ match(rownames(dat), gef$biosample_accession)]
    dat$study_time_collected_unit <- gef$study_time_collected_unit[ match(rownames(dat), gef$biosample_accession)]
    if(timeAs == "observation"){
      dat$sub_tp <- paste0( ptid, "_", dat$study_time_collected)
    }else{
      dat$participant_id <- ptid
    }
  }

  dat$stc_key <- paste(dat$study_time_collected, dat$study_time_collected_unit)
  dat <- dat[ dat$stc_key %in% timePts, ]
  if(nrow(dat) == 0){ return() } # Assay does not contain chosen timepoints
  
  if(assay %in% c("hai","neut_ab_titer")){
    sum_col <- paste0("mean_", assay)
    dat <- dat %>%
      group_by_("participant_id", "study_time_collected") %>%
      mutate(avg = mean(value_reported)) %>%
      summarize(sum_col = unique(avg))
    colnames(dat)[ colnames(dat) == "sum_col"] <- sum_col
    if(timeAs == "observation"){
      dat$sub_tp <- paste0( dat$participant_id, "_", dat$study_time_collected)
      dat <- dat[ , grep("sub|mean", colnames(dat))]
    }else{
      dat <- spread(dat, key = "study_time_collected", value = sum_col)
      nms <- colnames(dat)[ colnames(dat) != "participant_id"  ]
      nms <- paste0(nms, "_", sum_col)
      colnames(dat) <- c("participant_id", nms)
    }

  }else if(assay == "fcs_analyzed_result"){
    cols <- c("participant_id", "study_time_collected", "population_definition_reported", "population_cell_number")
    dat <- dat[ , cols, with = FALSE ]
    if(timeAs == "observation"){
      dat <- dat %>%
        mutate(sub_tp = paste0(participant_id, "_", study_time_collected)) %>%
        group_by(sub_tp, population_definition_reported) %>%
        summarize(newValue = unique(mean(population_cell_number)))
      dat <- spread(dat, key = "population_definition_reported", value = "newValue")
    }else{
      dat <- dat %>%
        mutate(popDef_day = paste(population_definition_reported, study_time_collected, sep = "_")) %>%
        group_by_("participant_id", "popDef_day") %>%
        summarise(newValue = unique(mean(population_cell_number)))
      dat <- spread(dat, key = "popDef_day", value = "newValue")
    }

  }else if(assay == "elisa"){
    cols <- c("participant_id", "study_time_collected", "analyte", "value_preferred")
    dat <- dat[ , cols, with = FALSE ]
    if(timeAs == "observation"){
      dat <- dat %>%
        mutate(sub_tp = paste0(participant_id, "_", study_time_collected)) %>%
        group_by(sub_tp, analyte) %>%
        summarize(newValue = unique(mean(value_preferred)))
      dat <- spread(dat, key = "analyte", value = "newValue")
    }else{
      dat <- dat %>%
        mutate(analyte_day = paste(analyte, study_time_collected, assay, sep = "_")) %>%
        group_by_("participant_id", "analyte_day") %>%
        summarise(newValue = unique(mean(value_preferred)))
      dat <- spread(dat, key = "analyte_day", value = "newValue")
    }

  }else if(assay == "elispot"){
    cols <- c("participant_id", "study_time_collected", "analyte", "spot_number_reported")
    dat <- dat[ , cols, with = FALSE ]
    if(timeAs == "observation"){
      dat <- dat %>%
        mutate(sub_tp = paste0(participant_id, "_", study_time_collected)) %>%
        group_by(sub_tp, analyte) %>%
        summarize(newValue = unique(mean(spot_number_reported)))
      dat <- spread(dat, key = "analyte", value = "newValue")
    }else{
      dat <- dat %>%
        mutate(analyte_day = paste(analyte, study_time_collected, assay, sep = "_")) %>%
        group_by_("participant_id", "analyte_day") %>%
        summarise(newValue = unique(mean(spot_number_reported)))
      dat <- spread(dat, key = "analyte_day", value = "newValue")
    }
  }else if(assay == "gene_expression_files"){
    # Be sure to only subset by GE timepoints not all assay time points!
    geOnlyTp <- unique(dat$stc_key)

    # Figure out subs with all tp
    subSums <- sapply(unique(dat$participant_id), function(x){
      return(length(dat$stc_key[ dat$participant_id == x]))
    })
    goodSubs <- names(subSums)[ subSums == length(geOnlyTp)]
    
    # Reorganize matrices by gs_timePt as colnames and cbind
    if(timeAs == "variable"){
      mats <- lapply(geOnlyTp, function(x){
        mat <- dat[ dat$stc_key == x & dat$participant_id %in% goodSubs, ]
        mat <- mat[, grep("stc_key|study_time_collected", colnames(dat), invert = T)]
        nms <- colnames(mat)[ colnames(mat) != "participant_id" ]
        nms <- paste0(nms, "_", gsub(" Days", "", x))
        colnames(mat) <- c(nms, "participant_id")
        mat <- mat[ order(match(mat$participant_id, goodSubs)), ]
        mat <- mat[ , -grep("participant_id", colnames(mat)) ]
        return(mat)
      })
      dat <- Reduce(function(x,y){cbind(x,y)}, mats)
      dat$participant_id <- goodSubs
    }else{
      dat <- dat[, grep("stc_key|study_time_collected", colnames(dat), invert = T)]
    }

  }else if(assay == "mbaa"){
    cols <- c("participant_id", "study_time_collected", "analyte", "concentration_value")
    dat <- dat[ , cols, with = FALSE ]
    if(timeAs == "observation"){
      dat <- dat %>%
        mutate(sub_tp = paste0(participant_id, "_", study_time_collected)) %>%
        group_by(sub_tp, analyte) %>%
        summarize(newValue = mean(as.numeric(concentration_value)))
      dat <- spread(dat, key = "analyte", value = "newValue")
    }else{
      dat <- dat %>%
        mutate(analyte_day = paste0(analyte, "_", study_time_collected)) %>%
        group_by(participant_id, analyte_day) %>%
        summarize(newValue = mean(as.numeric(concentration_value)))
      dat <- spread(dat, key = "analyte_day", value = "newValue")
    }

  }else if(assay == "pcr"){
    cols <- c("participant_id", "study_time_collected", "gene_symbol", "value_reported")
    dat <- dat[ , cols, with = FALSE ]
    if(timeAs == "observation"){
      dat <- dat %>%
        mutate(sub_tp = paste0(participant_id, "_", study_time_collected)) %>%
        group_by(sub_tp, gene_symbol) %>%
        summarize(newValue = mean(value_reported))
      dat <- spread(dat, key = "gene_symbol", value = "newValue")
    }else{
      dat <- dat %>%
        mutate(gs_day = paste0(gene_symbol, "_", study_time_collected)) %>%
        group_by(participant_id, gs_day) %>%
        summarize(newValue = mean(value_reported))
      dat <- spread(dat, key = "gs_day", value = "newValue")
    }
  }
  return(data.frame(dat, stringsAsFactors = FALSE))
}
```

```{r prep-assay-data}
# Sdy == "" if at project level!
con <- CreateConnection(sdy, onTest = grepl("test", baseUrl))

# get assay data
# NOTES:
# 1. When timeAs == "variable", GE data is subset to subjects with all timepoints
# Otherwise too much imputation is done and would affect subjects score.
# 2. If impute == "none", then assay data matrix is cut down to only subjects with
# every assay * timepoint combination. This could be highly limiting in project level work.
aData <- lapply(assays,
                getAssayData,
                con = con,
                timePts = timePts,
                timeAs = timeAs)

# Check aData to be sure non-null
nulls <- sapply(aData, is.null)
aData <- aData[ !nulls ]

# Impute missing values within each assay on per sample or sample_tp (column) basis if selected by user
if(impute == "mean"){
  aData <- lapply(aData, function(df){ 
    df <- apply(df, 2, function(x){ x[is.na(x) ] <- mean(x, na.rm = T); return(x)})
  })
}else if(impute == "median"){
  aData <- lapply(aData, function(df){ 
    df <- apply(df, 2, function(x){ x[is.na(x) ] <- median(x, na.rm = T); return(x)})
  })
}else if(impute == "knn"){
  aData <- lapply(aData, function(df){ 
    res <- impute::impute.knn(as.matrix(df))
    return(res$data)
  })
}else if(impute == "none"){
  aData <- lapply(aData, function(df){ 
    df <- df[ complete.cases(df), ]
  })
}

# Find subs with all datasets
subs <- sapply(aData, function(x){
  x[, grep("sub_tp|participant_id", colnames(x))]
})
numSubs <- table(unlist(subs))
subs <- names(numSubs)[ numSubs == length(aData) ]

# Remove subs without all datasets and order for cbind
aData <- lapply(aData, function(x){
  nmsCol <- grep("sub_tp|participant_id", colnames(x), value = T)
  x <- x[ x[[nmsCol]] %in% subs, ]
  x <- x[ order(match(x[[nmsCol]], subs)), ]
  x <- x[, -grep(nmsCol, colnames(x)) ]
})

# combine
allMat <- Reduce(function(x,y){cbind(x,y)}, aData)
rownames(allMat) <- subs

# normalize
allMat <- scale(allMat)

if( is.null(dim(allMat)) || dim(allMat)[2] < 2){
  stop("R Report Error: Too few variables selected. Please select more assays or timepoints and re-run.")
}else if(dim(allMat)[1] < 1){
  stop("R Report Error: No subjects with complete cases. Select imputation method other than 'None' and retry.")
}
```


```{r prep-demo-data}
# Get demographic data and add cohort name
demo <- con$getDataset("demographics")
cohortMap <- Rlabkey::labkey.selectRows(baseUrl = con$config$labkey.url.base,
                                        folderPath = con$config$labkey.url.path,
                                        schemaName = "study",
                                        queryName = "cohort",
                                        showHidden = TRUE)
demo$cohort <- cohortMap$Label[ match(demo$cohort, cohortMap$Rowid) ]
```

## Parameters
- **Timepoints**: `r timePts`
- **Assays**: `r assays`
- **Time As**: `r timeAs`
- **Plot Type**: `r plotType`
- **Number of Components Analyzed**: `r numComponents`
- **Imputation Method**: `r impute`
- **Immune Response Label**: `r responseVar`

## General Information
Note: Only subjects with assay data at all timepoints are included in the final
dimension reduction analysis. By default, there is no imputation. If you do select
to impute data, it is done within the individual assay data prior to combining multiple
assays.  For gene-expression data, imputation only occurs if subjects have at least
partial data for the timepoint. For non-gene-expression data, imputation does not require
partial data for the timepoint (e.g. some hai strains but not others).

- **Number of Observations (Subjects)**: `r dim(allMat)[1]`
- **Number of Features (Assay Data)**: `r dim(allMat)[2]`

```{r plot helpers}
# helper for iterating through multiple components
genPlot <- function(i, j, resObj, type, prettyLbl, colorCol){
  if(type == "tSNE"){
    df <- data.frame(x = resObj$Y[,i],
                     y = resObj$Y[,j],
                     stringsAsFactors = FALSE)
  }else{
    df <- data.frame(x = resObj$x[, paste0("PC",i)],
                     y = resObj$x[, paste0("PC",j)],
                     stringsAsFactors = FALSE)
  }

  p <- plot_ly(data = df,
               x = ~x,
               y = ~y,
               color = colorCol,
               hoverinfo = 'text',
               text = paste(prettyLbl, ": ", colorCol, "<br>ID: ", subs)) %>%
    layout(xaxis = list(title = paste0(type, i)),
           yaxis = list(title = paste0(type, j)),
           annotations = list(yref = 'paper',
                              xref = 'paper',
                              y = 1.02,
                              x = 1.175,
                              text = prettyLbl,
                              showarrow = FALSE),
           autosize = T,
           width = 700,
           showlegend=TRUE)
}

makePlot <- function(label, timeAs, demo, numComponents, plotType, allMat){
  prettyLbl <- tools::toTitleCase(gsub("_", " ", label))

  if(label == "time"){
    colorCol <- gsub("^SUB\\d{5,6}\\.\\d{2,4}_", "", subs)
  }else if(label == "study"){
    colorCol <- gsub("^SUB\\d{5,6}\\.", "", subs)
    colorCol <- paste0("SDY", gsub("_\\d+$", "", colorCol))
  }else if(label %in% c("hai", "neut_ab_titer")){
    re <- re[participant_id %in% subs ]
    # create response value from fold-change compared to baseline for each virus
    re <- re[, list(arm_accession, study_time_collected, study_time_collected_unit,
                    response = value_reported/mean(value_reported[study_time_collected<=0])),
             by = "virus,participant_id"]
    # determine mean response for each study day across virus
    re <- re[, mr := mean(response, na.rm=T), by = "study_time_collected"]
    # determine maximum mean response for each cohort for each study day
    re <- re[, ma := max(mr, na.rm=T), by = "arm_accession"]
    # Determine the peak study day for each cohort
    peak <- unique(re[mr == ma, list(study_time_collected, arm_accession)])
    # Use merge to filter to study days for peak response by cohort
    re <- merge(re, peak, by=c("study_time_collected", "arm_accession"))
    # provide single response for each participant by taking the log of max
    re <- re[, list(response = log2(max(response))), by = "participant_id"]
    # Dichotomize the participants into high and low responders arbitrarily at log2(4)
    # based on default values in IRP
    re <- re[, response := ifelse(response >= log2(4), "high responder",  "low responder")]
    colorCol <- re$response[order(match(subs, re$participant_id))]
  }else{
    if(timeAs == "observation"){ subs <- gsub("_\\d{1}", "", subs) }
    colorCol <- demo[[label]][ match(subs, demo$participant_id)]
  }
  # create and sort possible component combinations
  poss <- expand.grid(j = 1:numComponents, i = 1:numComponents)
  poss <- poss[ poss$j != poss$i, ]
  poss$test <- sapply(1:nrow(poss), function(x){
    paste(sort(c(poss[x,1], poss[x,2])), collapse = ":")
  })
  poss <- poss[ !duplicated(poss$test), ]

  # Run analysis and get pieces together for plotting
  if(plotType == "tSNE"){
    set.seed(8)
    # Avoid duplicate rows by adding random noise
    if( any(duplicated(allMat)) ){
      allMat[,1] <- jitter(allMat[,1], factor = 1, amount = NULL)
    }
    tRes <- Rtsne::Rtsne(allMat, dims = numComponents, perplexity = perplexity)
    moreArgs <- list(resObj = tRes, type = "tSNE", prettyLbl = prettyLbl, colorCol = colorCol)
  }else{
    pca <- prcomp(allMat)
    moreArgs <- list(resObj = pca, type = "PCA", prettyLbl = prettyLbl, colorCol = colorCol)
  }

  plots <- mapply(genPlot,
                  i = poss$i,
                  j = poss$j,
                  MoreArgs = moreArgs,
                  SIMPLIFY = FALSE
  )
}
```

## Plots {.tabset}

### Age
```{r  age, fig.align="left"}
plots <- makePlot(label = "age_reported",
                  timeAs = timeAs,
                  demo = demo,
                  numComponents = numComponents,
                  plotType = plotType,
                  allMat = allMat
)
htmltools::tagList(plots)
```

### Gender
```{r  gender, fig.align="left"}
plots <- makePlot(label = "gender",
                  timeAs = timeAs,
                  demo = demo,
                  numComponents = numComponents,
                  plotType = plotType,
                  allMat = allMat
)
htmltools::tagList(plots)
```

### Ethnicity
```{r  ethnicity, fig.align="left"}
plots <- makePlot(label = "ethnicity",
                  timeAs = timeAs,
                  demo = demo,
                  numComponents = numComponents,
                  plotType = plotType,
                  allMat = allMat
)
htmltools::tagList(plots)
```

### Race
```{r  race, fig.align="left"}
plots <- makePlot(label = "race",
                  timeAs = timeAs,
                  demo = demo,
                  numComponents = numComponents,
                  plotType = plotType,
                  allMat = allMat
)
htmltools::tagList(plots)
```

### Cohort
```{r  cohort, fig.align="left"}
plots <- makePlot(label = "cohort",
                  timeAs = timeAs,
                  demo = demo,
                  numComponents = numComponents,
                  plotType = plotType,
                  allMat = allMat
)
htmltools::tagList(plots)
```

### Timepoint
```{r  timePt, fig.align="left"}
if(timeAs == "observation"){
  plots <- makePlot(label = "time",
                    timeAs = timeAs,
                    demo = demo,
                    numComponents = numComponents,
                    plotType = plotType,
                    allMat = allMat
  )
  htmltools::tagList(plots)
}else{
  cat("Plots are only labelled by timepoint if `Use Time As` equals 'observation'.")
}
```

### Study
```{r  byStudy, fig.align="left"}
if(sdy == "/Studies"){
  plots <- makePlot(label = "study",
                    timeAs = timeAs,
                    demo = demo,
                    numComponents = numComponents,
                    plotType = plotType,
                    allMat = allMat
  )
  htmltools::tagList(plots)
}else{
  cat("Plots are only labelled by study if run at the project level through `Data Finder`.")
}
```

### Immune Response
Note: Not all subjects may have immune response data
```{r IR, fig.align="left"}
re <- tryCatch(
  con$getDataset(responseVar, original_view = T),
  warning = function(w) return(w),
  error = function(e) return(e)
)
if(is.null(re$message)){
  plots <- makePlot(label = responseVar,
                    timeAs = timeAs,
                    demo = demo,
                    numComponents = numComponents,
                    plotType = plotType,
                    allMat = allMat
  )
  htmltools::tagList(plots)
}else{
  cat(paste0("Response variable ", responseVar, " returns empty data frame from study.  Data is not available for given participants."))
}
```


<script>
window.HTMLWidgets.staticRender();
</script>

