---
output:
    html_document:
        df_print: paged
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
suppressPackageStartupMessages({
  library(knitr)
  library(magrittr)
  library(data.table)
})
getFileUrl <- function(filepath) {
  sub("/share/files", "/_webdav", filepath)
}
getFileLink <- function(filepath, filename) {
  paste0("<a href=\"", getFileUrl(filepath), "\">", filename, "</a>")
}
```


## All Data

<a href="/_webdav/HIPC/IS2/@files/ImmuneSignatures2.zip"><button>Download ImmuneSignatures2.zip (`r utils:::format.object_size(file.size("/share/files/HIPC/IS2/@files/ImmuneSignatures2.zip"), "auto")`)</button></a>

__DOI:__ [10.6084/m9.figshare.17096978](https://doi.org/10.6084/m9.figshare.17096978.v1)

__Note:__ The ImmuneSignatures data resource is derived from publicly available data on ImmPort, and available for reuse under the ImmPort [data use agreement](https://www.immport.org/agreement).

__Data Format__: Most of the datasets are presented here as `.rds` files, that is,
R objects which can be read into an R session via `readRDS()`


## Expression Sets

<div class="is2-downloads table-condensed">
```{r downloads, echo = FALSE}
files <- fread("/share/files/HIPC/IS2/@files/data/dataset_metadata.csv")
esets <- files[grepl("_eset\\.rds", dataset)]
eset_descriptions = c(
  all_noNorm_eset.rds = "All subjects, log2-normalized expression",
  all_noNorm_withResponse_eset.rds = "All subjects with immune response data, log2-normalized expression",
  all_norm_eset.rds = "All subjects, cross-study normalized and batch corrected expression",
  all_norm_withResponse_eset.rds = "All subjects with immune response data, cross-study normalized and batch corrected expression",
  young_noNorm_eset.rds = "Subjects age 18-50, log2-normalized expression",
  young_noNorm_withResponse_eset.rds = "Subjects age 18-50 with immune response data, log2-normalized expression",
  young_norm_eset.rds = "Subjects age 18-50, cross-study normalized and batch corrected expression",
  young_norm_withResponse_eset.rds = "Subjects age 18-50 with immune response data, cross-study normalized and batch corrected expression",
  old_noNorm_eset.rds = "Subjects age 60-90, log2-normalized expression",
  old_noNorm_withResponse_eset.rds = "Subjects age 60-90 with immune response data, log2-normalized expression",
  old_norm_batchCorrectedFromYoung_eset.rds ="Subjects age 60-90, cross-study normalized and batch corrected using correction coefficients from young",
  old_norm_batchCorrectedFromYoung_withResponse_eset.rds = "Subjects age 60-90 with immune response data, cross-study normalized and batch corrected using correction coefficients from young",
  extendedOld_noNorm_eset.rds = "Subjects age 50-90, log2-normalized expression",
  extendedOld_noNorm_withResponse_eset.rds = "Subjects age 50-90 with immune response data, log2-normalized coutns",
  extendedOld_norm_batchCorrectedFromYoung_eset.rds = "Subjects age 50-90, log2-normalized counts",
  extendedOld_norm_batchCorrectedFromYoung_withResponse_eset.rds = "Subjects age 50-90 with immune response data, cross-study normalized and batch corrected using correction coefficients from young"
)
esets <- esets[, description := eset_descriptions[dataset]]

esets <- esets[, .(
  File = as.character(getFileLink(path, dataset)),
  Description = as.character(description),
  Participants = subjects,
  Samples = samples,
  Features = features
)]

knitr::kable(esets, escape = FALSE, format = "html")
```
</div>


## Other Datasets

<div class="is2-downloads table-condensed">
```{r immdata, echo = FALSE, warning=FALSE}
immdata<- files[dataset == "all_immdata_with_response.rds"]
immdata$description <- "Immune response data for all subjects where available. This dataset includes one entry per participant, with metadata and the derived immune response call."
immdata_tbl <- immdata[, .(
  File = as.character(getFileLink(path, dataset)),
  Description = as.character(description),
  Participants = subjects,
  Samples = samples
)]
knitr::kable(immdata_tbl, escape = FALSE, format = "html")
```
</div>


---


## Code

Rmarkdown outputs from preprocessing, including code and some summaries. Preprocessing depends on the [ImmuneSignatures2](https://github.com/RGLab/ImmuneSignatures2) R package. (DOI: [10.5281/zenodo.5706261](https://zenodo.org/record/5706261#.YaVs1caIZQI))

1. [Pull Expressionsets](/_webdav/HIPC/IS2/%40files/data/html_outputs/pull_esets.html){target="_blank"}
    * Pulls gene expression data from ImmuneSpace server, constructs expressionsets,
    and saves output to an R object on disk.
1. [Generate Base ExpressionSet](/_webdav/HIPC/IS2/%40files/data/html_outputs/generate_base_eset.html){target="_blank"}
    * Perform basic QC and filtering, and save output to one ExpressionSet.
1. [Create Final ExpressionSets](/_webdav/HIPC/IS2/%40files/data/html_outputs/create_final_esets.html){target="_blank"}
    * Perform cross-study normalization and batch correction, and save the different
    versions available in the table above.


<style>
  .is2-downloads td, .is2-downloads th {
    border: 1px solid #ddd;
    padding: 3px;
  }
  .is2-downloads tr:nth-child(even){background-color: #f2f2f2;}
  .is2-downloads tr:hover {background-color: #ddd;}
  .is2-downloads th {
    background-color: #255c97;
    text-align: left;
    color: white;
    font-size: 15px;
    font-weight: bold;
    padding: 3px;
  }
</style>