
```{r knitr-options, echo=FALSE, eval=TRUE, cache=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4, dpi=100, fig.align="center")
options(width=80)

## Custom hooks to add caption to figures
knit_hooks$set(hfig.cap = function(before, options, envir) {
  if(!before & options$hfig.cap) {
    paste0('<div class="caption">',options$fig.cap,"</div>\n", sep="")
  }
})

## Custom hooks to add caption to figures
knit_hooks$set(htab.cap = function(before, options, envir) {
  if(before & options$htab.cap) {
    paste0('<div class="caption">',options$fig.cap,"</div>", sep="")
 } else if(options$htab.cap) {
    paste0("\n\n")
 }
})
```

```{r libraries, message=FALSE, echo=FALSE}
library(ggplot2)
library(reshape2)
library(data.table)
library(RColorBrewer)
library(Rlabkey)
library(Biobase)
library(pheatmap)
library(glmnet)
```
```{r plots-style, echo=FALSE}
mypalette <- rev(brewer.pal(name="PiYG", n=11))
brewer.qual <- "Paired"
IS_theme <- theme_bw(base_size=14)
```


```{r envir-var}
param <- list(
    timepoint   = as.numeric( labkey.url.params$timePoint ),
    GEA_acc     = labkey.url.params$analysisAccession,
    EM_acc      = labkey.url.params$expressionMatrixAccession,
    dichValue   = as.numeric( labkey.url.params$dichotomizeValue ),
    FDR_thresh  = as.numeric( labkey.url.params$fdrThreshold ),
    FC_thresh   = as.numeric( labkey.url.params$fcThreshold ),
    dichotomize = as.logical( labkey.url.params$dichotomize ),
    EM_pred     = labkey.url.params$expressionMatrixPredict
)
```

```{r eset-func, cache=FALSE}

eset_from_EM <- function(EM){
  em_filter <- makeFilter(c("expression_matrix_accession", "IN", paste(EM, collapse=";")))
  exMat <- labkey.selectRows(baseUrl = labkey.url.base, folderPath = labkey.url.path, schemaName = "study",
                             colFilter=em_filter, queryName = "gene_expression_matrices", colNameOpt = "rname")
  rownames(exMat) <- exMat$biosample_accession
  umat <- unique(exMat[, c("expression_matrix_accession", "download_link_download_link", "feature_mapping_file", "feature_mapping_file_link_download_link", "matrix_description_description")])
  EM_files <- gsub(".*%40files", labkey.file.root, umat$download_link_download_link)
  #exprs <- as.matrix(read.table(EM_files, sep="\t"))
  header <- scan(EM_files, what="character", nlines=1, sep="\t", quiet=TRUE)
  exprs <- fread(EM_files)
  exprs <- data.frame(exprs[, 2:ncol(exprs), with=FALSE], row.names=exprs[, 1, with=FALSE][[1]])
  colnames(exprs) <- header
  exprs <- as.matrix(exprs)

  fdata_links <- gsub(".*%40files", labkey.file.root, umat[[grep("feature.*download", colnames(umat))]])
  fdata <- read.table(fdata_links, header=TRUE, sep="\t")
  fdata <- fdata[na.omit(match(rownames(exprs), fdata$feature_id)),]
  phenodata <- exMat[ exMat$expression_matrix_accession==EM,]
  phenodata <- phenodata[ match(colnames(exprs), phenodata$biosample_accession),]
  phenodata <- as(phenodata, "AnnotatedDataFrame")
  fdata <- as(fdata, "AnnotatedDataFrame")
  eset <- ExpressionSet(assayData = exprs, phenoData = phenodata, featureData = fdata)
  return(eset)
  }
```



```{r get-data}
s2 <- system.time({
GEM <- eset_from_EM(param$EM_acc)
GER_filter <- makeFilter(c("analysis_accession", "EQUAL", param$GEA_acc),
                         c("adj_p_val", "LESS_THAN", param$FDR_thresh))
ds_GER <- labkey.selectRows(baseUrl = labkey.url.base, folderPath = labkey.url.path, schemaName = "lists", queryName = "gene_expression_analysis_result_wide", colNameOpt="rname", colFilter=GER_filter)
GER <- data.table(ds_GER)

})
```

```{r subset, echo=FALSE}
GER <- GER[abs(log_fc) > param$FC_thresh, ]
if(nrow(GER)==0){
  #GER_filter <- makeFilter(c("analysis_accession", "EQUAL", param$GEA_acc))
  #ds_GER <- labkey.selectRows(baseUrl = labkey.url.base, folderPath = labkey.url.path, schemaName = "lists", queryName = "gene_expression_analysis_result_wide", colNameOpt="rname", colFilter=GER_filter)
  #maxFC <- max(abs(ds_GER$log_fc))
  #minFDR <- min(ds_GER$adj_p_val)
  opts_chunk$set(cache=FALSE, echo=FALSE, eval=FALSE)
  stop("There are 0 probes selected. Try lowering the FDR threshold or increasing the absolute log-FoldChange threshold.")
       #"The minimum FDR for this cohort at this timepoint is ", minFDR,". Selected:", param$FDR_thresh, "\n",
       #"The maximum log Fold-Change is ", maxFC,". Selected:", param$FC_thresh)
}
probes_selected <- unique(GER$feature_id)

all_tp <- function(eset, timepoint){
  pd <- data.table(pData(eset))
  timepoints <- c(0, timepoint)
  pd <- pd[,keep:=(sum(study_time_reported%in%timepoints)==length(timepoints)), by="biosample_accession_name,subject_accession"]
  pd <- pd[keep==TRUE]
  pd <- pd[study_time_reported %in% timepoints]
  return(pd)
}
pd_GEM <- all_tp(GEM, param$timepoint)
GEM <- GEM[probes_selected, pd_GEM$biosample_accession] #sorted by subject, time
```

```{r lFC}
lFC <- exprs(GEM)[, GEM$study_time_reported == param$timepoint] - exprs(GEM)[, GEM$study_time_reported == 0]
```

```{r hai}
process_hai <- function(eset, dichotomize){ 
  HAI_filter <- makeFilter(c("subject_accession", "IN", paste(eset$subject_accession, collapse=";")))
  ds_HAI <- labkey.selectRows(baseUrl=labkey.url.base, schemaName="study", folderPath=labkey.url.path, queryName ="hai", colNameOpt="rname", colFilter=HAI_filter)
  HAI <- data.table(ds_HAI)
  HAI <- HAI[subject_accession %in% eset$subject_accession & biosample_accession_name %in% eset$biosample_accession_name]
  
  HAI <- HAI[,list(study_time_collected=study_time_collected, response=value_reported/value_reported[study_time_collected==0]),by="virus_strain,biosample_accession_name,subject_accession"]
  HAI <- HAI[study_time_collected==28]
  HAI <- HAI[,list(response=log2(max(response))),by="subject_accession"]
  if(isTRUE(dichotomize)){
    HAI <- HAI[,response:=ifelse(response>=log2(dichValue), FALSE,  TRUE)]
  }
  return(HAI)
}
OUTCOME_train <- process_hai(GEM, dichotomize=param$dichotomize)
```

```{r glmnet}
#glmnet
x <- t(lFC)
y <- OUTCOME_train$response
fit_hai <- glmnet(x, y, alpha=0.5)
cv_fit_hai <- cv.glmnet(x, y)
coef_hai <- predict(fit_hai, s=cv_fit_hai$lambda.min, type="coefficients")
```

```{r probes_selection}
#Selected probes
selected_probes <- names(which(abs(coef_hai[,1]) > 0))
selected_probes <- fData(GEM)[selected_probes[grep("Intercept", selected_probes, invert=TRUE)],]
#If the user selected less variables than observation, no selection is done
if(length(probes_selected) < nrow(OUTCOME_train)){
  x.selected <- x
  selected_probes <- fData(GEM)[probes_selected,]
} else{
  x.selected <- x[, colnames(x) %in% selected_probes$feature_id, drop=FALSE]
}

if(nrow(selected_probes) < 2){
  opts_chunk$set(eval=FALSE, cache=FALSE)
  stop("No probes were selected as predictive. You may try to change the filtering criteria.")
}
```

### Probe selection via the Elastic Net
The model selection is based on `r length(probes_selected)` selected probes. Optimal parameters for the Elastic Net are selected via cross-validation.
The best predictive probes are summarized in Table 1, along with t-values and p-values calculated by fitting a multivariate linear model with the selected probes.
```{r lasso}
#lasso
nFeatures <- dim(x.selected)[2]
nObs <- length(y)
if(nObs < nFeatures){
  cat("You selected more features (", nFeatures, ") than observations (", nObs, ").\nThe ", nObs-2, "most significant features will be kept.\n")
  topFeatures <- sort(abs(coef_hai[coef_hai[,1] >0,]), decreasing=TRUE)
  topFeatures <- topFeatures[grep("Intercept", names(topFeatures), invert=TRUE)]
  topFeatures <- topFeatures[1:(nObs-2)]
  x.selected <- x[,names(topFeatures)]
  selected_probes <- fData(GEM)[names(topFeatures),]
}
relasso <- lm(y~x.selected)
sum_relasso <- summary(relasso)
sum_relasso_coef <- sum_relasso$coefficients
predictor_table <- cbind(selected_probes, sum_relasso_coef[,c("t value","Pr(>|t|)")][-1,])
setnames(predictor_table, c("t value","Pr(>|t|)"), c("t-value","p-value"))
```

```{r table}
#Display
#Table
predictor_table[,"gene_symbol"] <- paste0('<a href="http://www.genenames.org/cgi-bin/quick_search.pl?.cgifields=type&type=equal&num=50&search=', gsub(";","+",predictor_table[,"gene_symbol"]),'&submit=Submit" target="_blank">', predictor_table[,"gene_symbol"], '</a>')
```

```{r kable, echo=FALSE, results='asis', htab.cap=TRUE, fig.cap="Table 1. Table of selected probes by the Elastic Net."}
kable(predictor_table, digits=2, format="html", row.names=FALSE, table.attr="id=\"res_table\"")
```

<br>
<h3> Heatmap </h3>
```{r heatmap, hfig.cap=TRUE, fig.cap="Figure 1. Heatmap of selected probes with HAI status annotation."}
#Heatmap
annotation <- data.frame(hai=as.numeric(OUTCOME_train$response))
rownames(annotation) <- colnames(lFC)
mat <- lFC[ as.character(predictor_table$feature_id), order(annotation$hai)]
pheatmap(mat, dendrogram="none", cluster_cols=FALSE, cluster_rows=TRUE,
         breaks = seq(-1, 1, length.out = length(mypalette) + 1),
         show_rownames=TRUE, show_colnames=FALSE, scale="none",cluster_method="ward",
         cluster_distance="correlation",color=mypalette, annotation=annotation,
         annotation_colors=list(hai=grey(10:0/10)))
```

```{r testing-cohort}
if(param$EM_pred != ""){
  GEP <- eset_from_EM(param$EM_pred)
  pd_GEP <- all_tp(GEP, param$timepoint)
  GEP <- GEP[probes_selected, pd_GEP$biosample_accession]
  lFC_pred <- exprs(GEP)[, GEP$study_time_reported == param$timepoint] - exprs(GEP)[, GEP$study_time_reported == 0]
  OUTCOME_pred <- process_hai(GEP, dichotomize=param$dichotomize)
}
```


### Predicting the response using the selected features
```{r predictionhfig.cap=TRUE, fig.cap="Figure 2. Predicted vs. observed values for the prediction model."}
#Prediction on training set
if(param$EM_pred ==""){
  df <- data.frame(observed=OUTCOME_train$response, fitted=relasso$fitted.values)
  ggplot(df, aes(x=observed, y=fitted)) + geom_point() + geom_smooth(method="lm") +IS_theme
} else{
  x.test <- data.frame(t(lFC_pred[ as.character(selected_probes$feature_id),]), check.names=FALSE)
  pred <- as.matrix(x.test)%*%as.vector((coef(relasso)[-1]))+coef(relasso)[1]
  df <- data.frame(observed = c(OUTCOME_train$response, OUTCOME_pred$response),
		   fitted = c(relasso$fitted.values, pred),
		   cohort = c(rep(paste("a:", param$EM_acc, "(Training)"), length(OUTCOME_train$response)),
			      rep(paste("a:", param$EM_pred, "(Testing)"), length(OUTCOME_pred$response))))
  ggplot(df, aes(x=observed, y=fitted)) + geom_point() + geom_smooth(method="lm") + IS_theme +
  	 facet_wrap(~cohort, scale="free") + xlab("Observed response") + ylab("Predicted response")
}
```

```{r printS, results="hide"}
param
labkey.url.params
```

