<link rel="stylesheet" type="text/css" href="https://ajax.aspnetcdn.com/ajax/jquery.dataTables/1.9.4/css/jquery.dataTables.css">


<style type="text/css">

#fw_container { width: 80%; margin: 0 auto; }

.caption {
   font-weight:bold;
   font-size:100%;
   text-align:center;
}

.knitr_table {
   margin: 0px auto; 
}

.knitr_table table { 
   border-collapse:collapse;   border-spacing:0px; margin:0px;
}
.knitr_table th { 
   padding:5px; font-weight:bold; background-color:#ccc;
}
.knitr_table td { 
   padding:5px;
}
.knitr_table tr:nth-child(odd) { 
   background-color:#fff; 
}
.knitr_table tr:nth-child(even) { 
   background-color:#eee; 
}
</style>

```{r knitr-options, echo=FALSE, eval=TRUE, cache=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4, dpi=100, fig.align="center")
options(width=80)

## Custom hooks to add caption to figures
knit_hooks$set(hfig.cap = function(before, options, envir) {
  if(!before & options$hfig.cap) {
    paste0('<div class="caption">',options$fig.cap,"</div>\n", sep="")
  }
})

## Custom hooks to add caption to figures
knit_hooks$set(htab.cap = function(before, options, envir) {
  if(before & options$htab.cap) {
    paste0('<div class="caption">',options$fig.cap,"</div>", sep="")
 } else if(options$htab.cap) {
    paste0("\n\n")
 }
})
```

```{r libraries, message=FALSE}
library(ggplot2)
library(reshape2)
library(data.table)
library(RColorBrewer)
library(Rlabkey)
library(Biobase)
library(pheatmap)
library(glmnet)
library(ISpaceR)
```
```{r plots-style}
mypalette <- rev(brewer.pal(name="PiYG", n=11))
brewer.qual <- "Paired"
IS_theme <- theme_bw(base_size=14)
```


```{r envir-var}
labkey.url.base <- "https://posey.fhcrc.org/"
labkey.url.path <- "/Emory/SDY61/"
labkey.url.param <-list(day = 7, GEA_acc = "GEA1", EM_acc = "EXPM0001", FDR_thresh = 0.02, FC_thresh = 0)
```

```{r reading-the tables, cache=TRUE}

GEM <- read_exprs_mat(param$EM_acc)[[1]]

GER_filter <- makeFilter(c("analysis_accession", "EQUAL", param$GEA_acc),
                         c("adj_p_val", "LESS_THAN", param$FDR_thresh))
ds_GER <- labkey.selectRows(baseUrl = labkey.url.base, folderPath = labkey.url.path, schemaName = "lists", queryName = "gene_expression_analysis_result_wide", colNameOpt="rname", colFilter=GER_filter)
GER <- data.table(ds_GER)

HAI_filter <- makeFilter(c("subject_accession", "IN", paste(GEM$subject_accession, collapse=";")))
ds_HAI <- labkey.selectRows(baseUrl=labkey.url.base, schemaName="study", folderPath=labkey.url.path, queryName ="hai", colNameOpt="rname", colFilter=HAI_filter)
HAI <- data.table(ds_HAI)
```

```{r subset}
#Filter-subset
#Filter out subjects with missing timepoints
#Filter selected probes
pd <- data.table(pData(GEM))
timepoints <- c(0, param$timepoint)
pd <- pd[,keep:=(sum(study_time_reported%in%timepoints)==length(timepoints)), by="biosample_accession_name,subject_accession"]
pd <- pd[keep==TRUE]
pd <- pd[study_time_reported %in% timepoints]
probes_selected <- unique(GER$feature_id)
GES <- GEM[probes_selected, pd$biosample_accession] #sorted by subject, time
HAI <- HAI[subject_accession %in% GES$subject_accession & biosample_accession_name %in% GES$biosample_accession_name]
```

```{r lFC}
#lFC
lFC <- exprs(GES)[, GES$study_time_reported == param$timepoint] - exprs(GES)[, GES$study_time_reported == 0]
```

```{r hai}
#hai
#Class subjects as high/low responders
HAI <- HAI[,list(study_time_collected=study_time_collected, hai_response=value_reported/value_reported[study_time_collected==0]),by="virus_strain,biosample_accession_name,subject_accession"]
HAI <- HAI[study_time_collected==28]
HAI <- HAI[,list(hai_response=max(hai_response)),by="subject_accession"]
HAI <- HAI[,responder:=ifelse(hai_response<=2, "Low", ifelse(hai_response>=4, "High", "Medium"))]
```

```{r glmnet}
#glmnet
x <- t(lFC)
y <- log2(HAI$hai_response)
fit_hai <- glmnet(x, y, alpha=0.5)
cv_fit_hai <- cv.glmnet(x, y)
coef_hai <- predict(fit_hai, s=cv_fit_hai$lambda.min, type="coefficients")
```

```{r probes_selection}
#Selected probes
selected_probes <- names(which(coef_hai[,1] > 0))
selected_probes <- fData(GES)[selected_probes[grep("Intercept", selected_probes, invert=TRUE)],]
#If the user selected less variables than observation, no selection is done
if(length(probes_selected) < nrow(HAI)){
  x.selected <- x
  selected_probes <- fData(GES)[probes_selected,]
} else{
  x.selected <- x[, colnames(x) %in% selected_probes$feature_id, drop=FALSE]
}

if(nrow(selected_probes) < 2){
  opts_chunk$set(eval=FALSE, cache=FALSE)
  stop("No probes were selected as predictive. You may try to change the filtering criteria.")
}
```

### Probe selection via the Elastic Net
The model selection is based on `r length(unique(probes_selected))` selected probes. Optimal parameters for the Elastic Net are selected via cross-validation.
The best predictive probes are summarized in Table 1, along with t-values and p-values calculated by fitting a multivariate linear model with the selected probes.
```{r lasso}
#lasso
relasso <- lm(y~x.selected)
sum_relasso <- summary(relasso)
sum_relasso_coef <- sum_relasso$coefficients
predictor_table <- cbind(selected_probes, sum_relasso_coef[,c("t value","Pr(>|t|)")][-1,])
setnames(predictor_table, c("t value","Pr(>|t|)"), c("t-value","p-value"))
```

```{r table}
#Display
#Table
predictor_table[,"gene_symbol"] <- paste0('<a href="http://www.genenames.org/cgi-bin/quick_search.pl?.cgifields=type&type=equal&num=50&search=', gsub(";","+",predictor_table[,"gene_symbol"]),'&submit=Submit" target="_blank">', predictor_table[,"gene_symbol"], '</a>')
```

```{r kable, echo=FALSE, results='asis', htab.cap=TRUE, fig.cap="Table 1. Table of selected probes by the Elastic Net."}
kable(predictor_table, digits=2, format="html", row.names=FALSE, table.attr="id=\"res_table\"")
```
<br>
<br>
<script type="text/javascript" charset="utf-8">
    $(document).ready(function() {
        $('#res_table').dataTable();
    } );
</script>
<br>
<br>

```{r echo=FALSE}
options(markdown.HTML.header = unlist(sapply(system.file("misc", c("vignette.css", "datatables.txt"), package = "knitr"), readLines)))
```

<h3> Heatmap of selected probes </h3>
```{r heatmap, hfig.cap=TRUE, fig.cap="Figure 1. Heatmap of selected probes with HAI status annotation."}
#Heatmap
annotation <- data.frame(hai=log2(HAI$hai_response))
rownames(annotation) <- colnames(lFC)
mat <- lFC[ as.character(predictor_table$feature_id), order(annotation$hai)]
pheatmap(mat, dendrogram="none", cluster_cols=FALSE, cluster_rows=TRUE, 
         breaks = seq(-1, 1, length.out = length(mypalette) + 1), 
         show_rownames=TRUE, show_colnames=FALSE, scale="none",cluster_method="ward",
         cluster_distance="correlation",color=mypalette, annotation=annotation, 
         annotation_colors=list(hai=grey(10:0/10)))
```

### Prediction on training set
```{r predictionhfig.cap=TRUE, fig.cap="Figure 2. Predicted vs. observed values for the prediction model."}
#Prediction on training set
x.test <- data.frame(t(lFC[ as.character(selected_probes$feature_id),]), check.names=FALSE)
hai_pred <- as.matrix(x.test)%*%as.vector((coef(relasso)[-1]))+coef(relasso)[1]
df <- data.frame(observed=HAI$hai_response, fitted=relasso$fitted.values)
ggplot(df, aes(x=log2(observed), y=fitted)) + geom_point() + 
  geom_smooth(method="lm") +IS_theme
```
