```{r dev, echo = FALSE, eval = FALSE}
#labkey.url.base <- "test.immunespace.org"
#labkey.url.path <- "Studies/SDY269"
#labkey.file.root <- "~/Downloads/"
```

```{r knitr-opts, echo = FALSE, message = FALSE, cache = FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE,
               fig.width=8, fig.height=4, dpi=100, fig.align="center",
               cache.path=paste(labkey.file.root, "cache/ImmuneResponsePredictor/", sep="/"))
```

```{r libraries, echo = FALSE, message = FALSE, cache = FALSE}
library(data.table)
library(Rlabkey)
co <- labkey.setCurlOptions(sslversion = 1, ssl.verifyhost = 2)
library(ggplot2)
library(glmnet)
library(pheatmap)
library(RColorBrewer)
library(RJSONIO)
library(ggthemr)
library(gplots)
library(ImmuneSpaceR)
library(Biobase)
```

```{r plot-styles, echo = FALSE, cache = TRUE}
palette <- ISpalette(20)
```

```{r envir-var, echo = FALSE, cache=FALSE}
param <- list(
  timepoint       = as.numeric(labkey.url.params$timePoint),
  timepointUnit   = labkey.url.params$timePointUnit,
  cohortsTraining = fromJSON(labkey.url.params$cohortsTraining),
  cohortsTesting  = fromJSON(labkey.url.params$cohortsTesting),
  GEA_acc         = fromJSON(labkey.url.params$analysisAccession),

  dichotomize     = as.logical(labkey.url.params$dichotomize),
  dichValue       = as.numeric(labkey.url.params$dichotomizeValue),
  FC_thresh       = as.numeric(labkey.url.params$fcThreshold),
  FDR_thresh      = as.logical(labkey.url.params$fdr)
)
```

```{r param, cache=TRUE, cache.extra=digest::digest(param)}
```

```{r functions, echo = FALSE, cache = TRUE}
get_GEAR <- function(GEA_acc, FC_thresh, FDR_thresh = FALSE){
  FDR_thresh <- ifelse(FDR_thresh, 0.1, 1) #No filtering if combobox isn't checked
  GEAR_filter <- makeFilter(c("analysis_accession", "IN", paste(GEA_acc, collapse=";")),
                            c("adj_p_val", "LESS_THAN", FDR_thresh),
                            c("log_fc", "NOT_BETWEEN", paste(-FC_thresh, FC_thresh, sep=",")))
  GEAR <-data.table(labkey.selectRows(baseUrl = labkey.url.base, folderPath = labkey.url.path, schemaName = "gene_expression", queryName = "gene_expression_analysis_results", colNameOpt="rname", colFilter=GEAR_filter))
  if(nrow(GEAR)==0){
    opts_chunk$set(cache=FALSE, echo=FALSE, eval=FALSE)
    stop("There are 0 probes selected. Try removing the the absolute log-FoldChange filtering under 'Additional options'.")
  }
  GEA_filter <- makeFilter(c("analysis_accession", "IN", paste(unique(GEAR$analysis_accession), collapse = ";")))
  GEA <- data.table(labkey.selectRows(baseUrl = labkey.url.base, folderPath = labkey.url.path, schemaName = "gene_expression", queryName = "gene_expression_analysis", colNameOpt = "rname", colFilter = GEA_filter))
  GEAR <- merge(GEAR, GEA, by = "analysis_accession")
  return(GEAR)
}

get_pdata <- function(EM_acc, timepoint){
  timepoints <- c(0, timepoint)
  pdata_filter <- makeFilter(c("Run/DataOutputs/Name", "IN", paste(EM_acc, collapse=";")))
  pdata <- unique(data.table(labkey.selectRows(labkey.url.base, labkey.url.path, 
		  "assay.ExpressionMatrix.matrix", "InputSamples", "gene_expression_matrices", 
		  colFilter = pdata_filter, colNameOpt = "rname")))
  setnames(pdata, colnames(pdata), gsub("^biosample_", "", colnames(pdata)))
  pdata <- pdata[, keep:=(sum(study_time_collected %in% timepoints) == length(timepoints)), by="subject_accession"]
  pdata <- pdata[keep == TRUE]
  pdata <- pdata[study_time_collected %in% timepoints]
  pdata <- pdata[order(subject_accession)]
  return(pdata)
}

get_DEG <- function(GEAR, pdata){
  em2fas <- unique(pdata[, list(run_dataoutputs_name, run_featureset)])
  FAS_filter <- makeFilter(c("FeatureAnnotationSetId", "IN", paste(em2fas$run_featureset, collapse = ";")))
  FAS <- data.table(labkey.selectRows(baseUrl = labkey.url.base, folderPath = labkey.url.path, schemaName="Microarray", queryName="FeatureAnnotation", colFilter = FAS_filter, colNameOpt = "rname", colSelect=c("featureid", "genesymbol")))
  setnames(FAS, c("featureid"), "feature_id")
  GEAR <- data.table(merge(GEAR, FAS, by = "feature_id"))
  DEG <- unique(GEAR$genesymbol) #rm duplicates and probes with no matches
  DEG <- DEG[!is.na(DEG)]
  return(DEG)
}

get_EM <- function(pdata, DEG){
  files <- file.path(labkey.file.root, "analysis/exprs_matrices", paste0(unique(pdata$run_dataoutputs_name), ".summary"))
  headers <- lapply(files, scan, what="character", nlines=1, sep="\t", quiet=TRUE)
  EMs <- lapply(files, fread)
  common_genes <- Reduce(intersect, lapply(EMs, "[",, "gene_symbol", with=FALSE))
  DEG <- DEG[DEG %in% common_genes$gene_symbol]
  EMs <- lapply(EMs, function(x){x[na.omit(match(DEG, x$gene_symbol))]})
  EMs <- lapply(EMs, function(x){t(x[ ,gene_symbol := NULL])})
  EM <- do.call("rbind", EMs)
  colnames(EM) <- DEG
  return(EM)
}

get_FC <- function(EM, pdata, timepoint){
  FC <- EM[pdata[study_time_collected == timepoint, biosample_accession], ] - EM[pdata[study_time_collected == 0, biosample_accession],]
  rownames(FC) <- pdata[match(rownames(FC), biosample_accession), subject_accession]
  return(FC)
}

get_HAI <- function(pdata, dichotomize, di_value){
  HAI_filter <- makeFilter(c("subject_accession", "IN", paste(pdata$subject_accession, collapse=";")))
  HAI <- data.table(labkey.selectRows(baseUrl=labkey.url.base, schemaName="study", folderPath=labkey.url.path, queryName ="hai", viewName = "full", colNameOpt="rname", colFilter=HAI_filter))
  HAI <- HAI[subject_accession %in% pdata$subject_accession]
  HAI <- HAI[,list(arm_accession, study_time_collected=study_time_collected, response=value_reported/value_reported[study_time_collected==0]),by="virus_strain,subject_accession"]
  HAI <- HAI[study_time_collected==28]
  HAI <- HAI[,list(response=log2(max(response))),by="subject_accession,arm_accession"]
  if(dichotomize){
    HAI <- HAI[,response:=ifelse(response>=log2(di_value), TRUE,  FALSE)]
  }
  cohorts <- unique(pdata[, list(arm_name, subject_accession)])
  HAI <- merge(HAI, cohorts, by = "subject_accession")
  HAI <- HAI[order(subject_accession)]
  return(HAI)
}

select_features <- function(FC, response, dichotomize){
  # Elastic net
  if(dichotomize){
    fit <- glmnet(FC, as.factor(response), alpha = 0.5, family = "binomial") 
  } else{
    fit <- glmnet(FC, response, alpha = 0.5)
  }
  cv_fit <- cv.glmnet(FC, response)
  coef <- predict(fit, s=cv_fit$lambda.min, type="coefficients")
  # Selecting features
  selected_features <- names(which(abs(coef[,1]) > 0))
  selected_features <- grep("Intercept", selected_features, invert = TRUE, value = TRUE)
  if(length(selected_features) < 2){
    opts_chunk$set(eval=FALSE, cache=FALSE)
    stop("No features were selected as predictive. You may try to remove the fold change filtering under 'Additional options'.")
  }
  # Need more obs than features
  nFeatures <- length(selected_features)
  nObs <- length(response)
  if(nObs <= nFeatures){
    cat("You selected as many or more features (", nFeatures, ") than observations (", nObs, ").\nThe", nObs-2, "most significant features will be kept.\n")
    selected_features <- names(sort(coef[selected_features,])[1:(nObs-2)])
  }
  return(selected_features)
}

# Heatmap #FC_selected, HAI, palette, dichotomize
draw_heatmap <- function(FC_selected, HAI, dichotomize, palette){
  annotation <- HAI[, list(subject_accession, response, arm_name)]
  annotation <- data.frame(annotation[order( response, arm_name, subject_accession)])
  rownames(annotation) <- annotation$subject_accession
  annotation$subject_accession <- NULL
  if(dichotomize){
    annotation$response <- as.factor(annotation$response)
    anno_col <- list(response=c(`FALSE`="white", `TRUE`="black"))
  } else{
    anno_col <- list(response=grey(10:0/10))
  }
  mat <- t(FC_selected)
  mat <- mat[, match(rownames(annotation), colnames(mat))]
  max_FC <- max(abs(mat))
  rownames(mat) <- ifelse(nchar(rownames(mat)) > 15, paste0(substr(rownames(mat), 1, 15), "..."), rownames(mat))
  pheatmap(mat, dendrogram = "none", cluster_cols = FALSE, cluster_rows = TRUE,
           breaks = seq(-max_FC, max_FC, length.out = length(palette) +1),
           color = palette, show_colnames = FALSE, scale = "row", 
           cluster_distance="correlation", cluster_method = "ward",
           annotation = annotation, annotation_colors = anno_col)
}

format_table <- function(x, ...){
  require(knitr)
  require(data.table)
  if(!is(x,"data.table")){
    #This is to make sure that we don't have issues with row names in data.frames
    stop("This function only works on data.tables")
  }
  out <- kable(x, format="html", digits=4, table.attr = 'class="display" id="res_table"', output = FALSE, ...)
  footer <- paste0("</thead>\n<tfoot>", strsplit(strsplit(out,"<thead>")[[1]][2],"</thead>")[[1]][1], "</tfoot>\n")
  out_new <- gsub("</thead>",footer,out)
  cat(out_new)
}
```

```{r funcs2, echo = FALSE}
getRun <- function(con, cohort){
  gem <- con$data_cache$GE_inputs
  run <- gsub(".tsv$", "", unique(gem[ gem$arm_name %in% cohort, "name"]))
  return(run)
}

# Get the hai response at peak immunogenicity for a subject list
# @value A data.table with response and subject_accession, ordered by subject (i.e: same order as FC or matrix)
getHAI <- function(con, subjects, dichotomize, dichValue){
  hai <- con$getDataset("hai", original_view = TRUE)
  hai <- hai[subject_accession %in% subjects]
  hai <- hai[, list(arm_accession, study_time_collected, study_time_collected_unit,
                    response = value_reported/mean(value_reported[study_time_collected<=0])),
               by = "virus_strain,subject_accession"]
  hai <- hai[, mr := mean(response), by="study_time_collected"]
  hai <- hai[, ma := max(mr), by = "arm_accession"]
  peak <- unique(hai[mr ==ma, list(study_time_collected, arm_accession)])
  hai <- merge(hai, peak, by=c("study_time_collected", "arm_accession"))
  #print(hai[, max(mr), by="arm_accession"])
  #immuno_peak <- hai[hai[, mr == max(mr)], study_time_collected]
  #
  #print(immuno_peak)
  #hai <- hai[study_time_collected == immuno_peak]
  hai <- hai[, list(response=log2(max(response))), by="subject_accession"]
  if(dichotomize){
    hai <- hai[,response:=ifelse(response>=log2(dichValue), TRUE,  FALSE)]
  }
  hai <- hai[order(subject_accession)]
  return(hai)
}
```


```{r no-probes, cache = FALSE, echo = FALSE}
if(param$timepoint > 0){
  GEAR <- get_GEAR(param$GEA_acc, param$FC_thresh, param$FDR_thresh)
}
```

```{r main, cache = FALSE}
con <- CreateConnection()
if(param$timepoint > 0){
  #EM <- con$getGEMatrix(
  pdata <- get_pdata(paste0(unique(GEAR$expression_matrix), ".tsv"), param$timepoint)
  DEG <- get_DEG(GEAR, pdata)
  EM <- get_EM(pdata, DEG)
  training_data <- get_FC(EM, pdata, param$timepoint)

  HAI <- get_HAI(pdata, param$dichotomize, param$dichValue)
  response <- HAI$response

} else{
  #$timepoint <= 0
  #$GEA_acc = NULL
  runs <- getRun(con, param$cohortsTraining)
  EM <- con$getGEMatrix(runs, summary = TRUE, reload = TRUE)
  EM <- EM[, pData(EM)$study_time_collected == param$timepoint & tolower(pData(EM)$study_time_collected_unit) == param$timepointUnit]
  subs <- pData(EM)$subject_accession
  HAI <- getHAI(con, subs, param$dichotomize, param$dichValue)
  HAI <- HAI[, arm_name := pData(EM)$arm_name]
  response <- HAI$response
  colnames(exprs(EM)) <- subs
  training_data <- t(exprs(EM))
}
```
```{r select-features, cache = FALSE}
selected_features <- select_features(training_data, response, param$dichotomize)

```{r lasso, cache = FALSE}
FC_selected <- training_data[,selected_features]

# lasso #IN FC_selected, dichotomize, selected_features #OUT predictor_table
FC_selected <- data.frame(FC_selected, check.names = FALSE)
colnames(FC_selected) <- gsub("@", "", gsub(";", "_OR_", gsub("-", "_", colnames(FC_selected)))) # TODO: No need for @ / _OR_
form <- as.formula(paste0("outcome~`", paste(colnames(FC_selected), collapse="`+`"), sep="`"))
FC_selected$outcome <- response
if(param$dichotomize){
     relasso <- glm(form, FC_selected, family="binomial")
} else{
     relasso <- lm(form, FC_selected)
}

FC_selected$outcome <- NULL
sum_relasso <- summary(relasso)
sum_relasso_coef <- sum_relasso$coefficients
pred_cIdx <- grep("value|Pr", colnames(sum_relasso_coef))
predictor_table <- sum_relasso_coef[,pred_cIdx][-1,]
colnames(predictor_table) <- c("statistic", "p-value")
predictor_table <- data.table(cbind(selected_features, predictor_table))
predictor_table <- predictor_table[, `:=`(c("statistic", "p-value"), list(as.numeric(statistic), as.numeric(`p-value`)))]
predictor_table <- predictor_table[, gene_symbol := selected_features]
# TODO: Fix links in DataTables
#predictor_table <- predictor_table[, gene_symbol := sapply(strsplit(as.character(selected_features), split=";"), function(X){ paste0('<a href="http://immunet-dev.princeton.edu/predictions/gene/?network=immune_global&gene=', X, '" target="_blank">', X, '</a>', collapse=";")})]
```

```{r pred-cohort}
if(length(param$cohortsTesting) > 0){
  run <- getRun(con, param$cohortsTesting)
  EM_pred <- con$getGEMatrix(run, summary = TRUE)
  EM_pred <- EM_pred[selected_features,]
  if(param$timepoint <= 0){
    EM_pred <- EM_pred[, EM_pred$study_time_collected == param$timepoint]
    EM_pred <- EM_pred[, pData(EM_pred)$study_time_collected == param$timepoint &
                         tolower(pData(EM_pred)$study_time_collected_unit) == param$timepointUnit]
    pdata <- data.table(pData(EM_pred))
    subs <- pdata$subject_accession
    data_pred <- t(exprs(EM_pred))
  } else{ #Need FC
    pdata <- data.table(pData(EM_pred))
    pdata <- pdata[study_time_collected %in% c(0, param$timepoint) & tolower(study_time_collected_unit) == param$timepointUnit]
    subs <- pdata[, .N, by = "subject_accession"][N>1, subject_accession]
    pdata <- pdata[subject_accession %in% subs]
    pdata <- pdata[order(subject_accession, study_time_collected)]
    data_pred <- exprs(EM_pred[, pdata[study_time_collected == param$timepoint, biosample_accession]]) -
          exprs(EM_pred[, pdata[study_time_collected == 0, biosample_accession]])
    data_pred <- t(data_pred)
    rownames(data_pred) <- pdata[match(rownames(data_pred), pdata$biosample_accession), subject_accession]
  }
  colnames(data_pred) <- colnames(FC_selected) #Handles the -/_@ issues
  HAI_pred <- getHAI(con, subs, param$dichotomize, param$dichValue)
  HAI_pred <- HAI_pred[, arm_name := pdata[, arm_name]]
}
```

```{r prediction, echo = FALSE, dev='CairoPNG'}
# Prediction
if(length(param$cohortsTesting) == 0){
     data <- data.frame(observed = response, fitted=relasso$fitted.values)
   p<- ggplot(data, aes(x=observed, y=fitted)) + geom_point() + geom_smooth(method="lm")
} else{
  data_pred <- data.frame(data_pred)
  predicted_val <- predict(relasso, newdata = data_pred, type = "response")
  data <- data.frame(observed = c(HAI$response, HAI_pred$response),
                     fitted = c(relasso$fitted.values, predicted_val),
                     cohort = c(paste("a:", HAI$arm_name, "(Training)"), paste("b:", HAI_pred$arm_name, "(Testing)")))
  p<- ggplot(data, aes(x=observed, y=fitted)) + geom_point() + geom_smooth(method="lm") +
    facet_wrap(~cohort, scale="free") + xlab("Observed response") + ylab("Predicted response")
}
  ggthemr('solarized')
  plot(p)
  
```
<br><br>

## Table of probes selected by the Elastic Net
```{r kable, results = 'asis'}
format_table(predictor_table)
```
<br><br>

## Heatmap of the selected features
```{r heatmap}
draw_heatmap(FC_selected, HAI, param$dichotomize, palette)
```
