```{r knitr, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, 
                      echo = FALSE, 
                      warning = FALSE, 
                      fig.align = 'left', 
                      fig.width = 12)
```

```{r libs, message = FALSE}
library(Rlabkey)
library(ImmuneSpaceR)
library(DT)
library(dplyr)
library(ggplot2)
library(plotly)
library(tidytext)
library(wordcloud)
library(lubridate)
library(readr)
library(stringr)
```

```{r preprocess}
#str(labkey.url.params)

# labkey parameters
from <- ifelse(labkey.url.params$from == "", as.Date("2016-01-01"), as.Date(labkey.url.params$from))
to <- ifelse(labkey.url.params$to == "", Sys.Date(), as.Date(labkey.url.params$to))
class(from) <- "Date"
class(to) <- "Date"
from <- ifelse(from > to, as.Date("2016-01-01"), from)
class(from) <- "Date"
from_to <- tibble(date = seq(from, to, by = "1 day"))

plotType <- ifelse(labkey.url.params$plotType %in% c("bar", "line"), labkey.url.params$plotType, "bar")
dateBy <- ifelse(labkey.url.params$by %in% c("day", "week", "month", "year"), labkey.url.params$by, "day")

# exclude admins and support accounts
admins <- labkey.selectRows(baseUrl = labkey.url.base,
                            folderPath = "/home", 
                            schemaName = "core", 
                            queryName = "SiteUsers", 
                            viewName = "", 
                            colFilter = makeFilter(c("Groups/Group$SName", "CONTAINS_ONE_OF", "Developers;Administrators")), 
                            containerFilter = NULL)
extras <- labkey.selectRows(baseUrl = labkey.url.base,
                            folderPath = "/home", 
                            schemaName = "core", 
                            queryName = "SiteUsers", 
                            viewName = "", 
                            colFilter = makeFilter(c("Email", "CONTAINS_ONE_OF", "rglab.org;immunespace.org")), 
                            containerFilter = NULL)
exclusion_list <- paste0(unique(c(admins$`Display Name`, extras$`Display Name`)), collapse = ";")
```



# Users
```{r newusers}
users_all <- labkey.selectRows(baseUrl = labkey.url.base,
                               folderPath = "/",
                               schemaName = "core",
                               queryName = "SiteUsers",
                               viewName = "",
                               colFilter = NULL,
                               containerFilter = NULL)
colnames(users_all) <- tolower(chartr(" ", "_", colnames(users_all)))
users_all <- users_all %>% 
  mutate(created = as.Date(created), 
         last_login = as.Date(last_login))

users <- users_all %>% 
  filter(created >= from & created <= to) %>% 
  select(created) %>% 
  count(created) %>% 
  right_join(from_to, by = c("created" = "date")) %>% 
  mutate(n = replace(n, is.na(n), 0)) %>% 
  mutate(created = floor_date(created, dateBy)) %>% 
  group_by(created) %>% 
  summarise(N = sum(n))
p <- ggplot(users, aes(x = created, y = N)) + 
  xlab(paste0("By ", dateBy)) + 
  ylab("New users") + 
  ggtitle("New users over time") + 
  scale_x_date(limits = c(from, to)) + 
  theme_IS(base_size = 18)
if (plotType == "line") {
  p <- p + geom_line()
} else {
  p <- p + geom_bar(stat = "identity")
}
ggplotly(p)
``` 

We gained **`r round(mean(users$N), 2)` new users** per *`r dateBy`* on average from `r from` to `r to`.
<br>


```{r activeusers}
logins <- labkey.selectRows(baseUrl = labkey.url.base,
                            folderPath = "/",
                            schemaName = "auditLog",
                            queryName = "UserAuditEvent",
                            viewName = "",
                            colFilter = makeFilter(c("CreatedBy/DisplayName", "NOT_IN", exclusion_list), 
                                                   c("Comment", "CONTAINS", "Database authentication")),
                            containerFilter= "AllFolders")
logins <- logins %>% 
  mutate(Date = as.Date(Date)) %>% 
  select(Date, `Created By`) %>% 
  filter(Date >= from & Date <= to)
active_by <- logins %>% 
  distinct() %>%
  count(Date) %>% 
  right_join(from_to, by = c("Date" = "date")) %>%
  mutate(n = replace(n, is.na(n), 0)) %>%
  mutate(Date = floor_date(Date, dateBy)) %>%
  group_by(Date) %>%
  summarise(n = sum(n))
p <- ggplot(active_by, aes(x = Date, y = n)) + 
  xlab(paste0("By ", dateBy)) + 
  ylab("Active users") + 
  ggtitle("Active users over time") + 
  scale_x_date(limits = c(from, to)) + 
  theme_IS(base_size = 18)
if (plotType == "line") {
  p <- p + geom_line()
} else {
  p <- p + geom_bar(stat = "identity")
}
ggplotly(p)
```

We had **`r round(mean(active_by$n), 2)` active users** per *`r dateBy`* on average from `r from` to `r to`.
<br>


```{r logins}
logins_by <- logins %>% 
  mutate(Date = floor_date(as.Date(Date), dateBy)) %>% 
  count(Date, `Created By`) %>% 
  left_join(select(users_all, user_id, display_name), by = c("Created By" =  "user_id")) %>% 
  mutate(display_name = ifelse(is.na(display_name), as.character(`Created By`), display_name)) %>% 
  select(Date, display_name, n)
if (nrow(logins_by) > 0) {
  p <- ggplot(logins_by, aes(x = Date, y = n, colour = display_name)) + 
    geom_line(alpha = 0.3) +
    xlab(paste0("By ", dateBy)) + 
    ylab("# of logins") + 
    ggtitle("Number of logins by users over time") + 
    scale_x_date(limits = c(from, to)) +
    theme(legend.position = "none") +
    theme_IS(base_size = 18)
  ggplotly(p)
}
```
<br>


```{r vs}
#vs <- full_join(df, logins_byday, by = c("created" = "Date")) %>% 
#  select(created, N, n) %>% 
#  mutate(N = ifelse(is.na(N), 0, N), n = ifelse(is.na(n), 0, n), surp = n - N)
#p <- ggplot(vs, aes(x = N, y = n)) + 
#  geom_point(aes(col = surp, label = created)) + 
#  geom_abline(intercept = 0) + 
#  theme_IS(base_size = 18)
#ggplotly(p)
```


```{r totalusers}
p <- users %>% 
  arrange(created) %>% 
  ggplot(aes(x = created, y = cumsum(N))) + 
  geom_step() + 
  xlab(paste0("By ", dateBy)) + 
  ylab("Total # of users") + 
  ggtitle("Total number of users over time") + 
  scale_x_date(limits = c(from, to)) + 
  theme_IS(base_size = 18)
ggplotly(p)
```
<br>


## Table of users
```{r usertable}
logins_byuser <- logins %>% 
  count(`Created By`)
topusers <- left_join(users_all, logins_byuser, by = c("user_id" = "Created By")) %>% 
  filter(!is.na(n)) %>% 
  arrange(desc(n))
datatable(topusers)
```
<br>



***



# Searches Queries
```{r searches}
searches <- labkey.selectRows(baseUrl = labkey.url.base, 
                              folderPath = "/home", 
                              schemaName = "auditLog", 
                              queryName = "SearchAuditEvent", 
                              viewName = "", 
                              colSort = "Created", 
                              colFilter = makeFilter(c("Query", "DOES_NOT_CONTAIN", "\\"), 
                                                     c("Query", "DOES_NOT_CONTAIN", "1234"), 
                                                     c("CreatedBy/DisplayName", "NOT_IN", exclusion_list)), 
                              containerFilter=NULL)

searches <- searches %>% 
  filter(`Created By` != 0, 
         Date >= from & Date <= to) %>% 
  mutate(id = paste(Date, `Created By`, sep = "-")) %>% 
  select(id, Query)
```

## Word cloud
```{r wordcloud}
freq <- searches %>% 
  unnest_tokens(output = word, input = Query) %>%
  count(word, sort = T)
pal <- brewer.pal(9,"YlGnBu")
pal <- pal[-(1:4)]

if (nrow(freq) > 0) {
  wordcloud(words = freq$word, 
            freq = freq$n,
            max.words = 30,
            rot.per = 0,
            random.order = FALSE,
            colors = pal)
} else {
  wordcloud(words = "no search", freq = 1)
}
```
<br>


## Table of search queries
```{r wordtable}
datatable(freq, width = 600)
```
<br>



***



# Studies and Modules
```{r studycount}
folders <- labkey.selectRows(baseUrl = labkey.url.base,
                             folderPath = "/",
                             schemaName = "auditLog",
                             queryName ="ContainerAuditEvent",
                             colNameOpt = "rname",
                             containerFilter = "AllFolders")
sdys <- folders %>% 
  filter(!is.na(container)) %>% 
  select(created, comment) %>% 
  filter(grepl("Folder SDY[0-9].*created$", comment)) %>% 
  mutate(study = gsub("\\s.*$", "", gsub("^.*SDY", "SDY", comment))) %>% 
  arrange(created) %>% 
  mutate(count = 1:nrow(.)) %>% 
  filter(created >= "2016-01-01") %>% 
  select(created, count) %>% 
  mutate(created = as.Date(created))
sdys <- bind_rows(tibble(created = as.Date("2016-01-01"), count = min(sdys$count)), 
                   sdys, 
                   tibble(created = Sys.Date(), count = max(sdys$count)))

# Based on folder creation -> counts private studies
p <- ggplot(sdys, aes(x = created)) +
  geom_step(aes(y = count)) +
  xlab("Date") + 
  ylab("Studies") + 
  ggtitle("Number of studies available") +
  scale_x_date() + 
  theme_IS(base_size = 18)
ggplotly(p)
```
<br>


```{r tomcat-logs}
# These logs are actually `/labkey/apps/tomcat/logs/` created by tomcat.
# Since they cannot be access by the r server, we need to copy them to `/share`
# Set up a cron job on `wsP/T` as `immunespace`
# crontab -e
# Add this line
# 00 0-23/6 * * * rsync -a -v /labkey/apps/tomcat/logs/localhost_access_log.* /share/tomcat-logs/
# This will sync logs to `/share/tomcat-logs/` every six hour.
read_log <- function(date) {
  if (Sys.info()["nodename"] == "immunetestrserve" & date > "2016-08-08") {
    file_name <- paste0("/share/tomcat-logs/localhost_access_log..", date, ".txt")
  } else {
    file_name <- paste0("/share/tomcat-logs/localhost_access_log.", date, ".txt")
  }
  if (file.exists(file_name)) {
    read_delim(file = file_name, 
               delim = " - ",
               escape_double = TRUE,
               col_names = FALSE,
               trim_ws = TRUE) %>% 
      mutate(date = date, 
             study = str_extract(X6, "SDY\\d+"), 
             schema = grepl("schemaName=study?", X14), 
             query = str_extract(X14, "(?<=queryName=)\\w+"))
  } else {
    NULL
  }
}
logs_dt <- lapply(from_to$date, read_log) %>% bind_rows()
```


## Pageviews: Study
```{r study-views}
log_study <- logs_dt %>% 
  mutate(folder = str_extract(X6, "(/project/Studies/SDY\\d+/begin.view\\?? |/project/Studies/begin.view\\?? )")) %>% 
  filter(!is.na(folder)) %>% 
  mutate(study = ifelse(is.na(study), "Data Finder", study)) %>% 
  select(study) %>% 
  count(study) %>% 
  arrange(desc(n))
datatable(log_study, 
          colnames = c("Study", "Unique pageviews"), 
          caption = paste0("From ", from, " to ", to), 
          width = 600)
```


## Pageviews: Module
```{r module-views}
log_module <- logs_dt %>% 
  mutate(module = str_extract(X6, "(?<=GET /)(DataExplorer|GeneExpressionExplorer|GeneSetEnrichmentAnalysis|ImmuneResponsePredictor)(?=/\\w+)")) %>% 
  filter(!is.na(module)) %>% 
  select(module) %>% 
  count(module) %>% 
  arrange(desc(n))
datatable(log_module, 
          colnames = c("Module", "Unique pageviews"), 
          caption = paste0("From ", from, " to ", to), 
          width = 600)
```


***



# ImmuneSpaceR

```{r isr}
ISR <- logs_dt %>% 
  filter(grepl("ImmuneSpaceR", X12)) %>% 
  mutate(study = ifelse(is.na(study), "All", study))
ISR_inits <- ISR %>% 
  filter(query == "ISC_study_datasets")
```

## Connections

```{r isr-study}
ISR_study <- ISR_inits %>%
  select(study) %>%
  count(study) %>%
  arrange(desc(n))
datatable(ISR_study,
          colnames = c("Study", "ISR connections"),
          caption = paste0("From ", from, " to ", to), 
          width = 600)
```
<br>


```{r isr-con}
ISR_con <- ISR_inits %>% 
  select(date) %>% 
  count(date) %>% 
  right_join(from_to, by = "date") %>% 
  mutate(n = replace(n, is.na(n), 0)) %>% 
  mutate(date = floor_date(date, dateBy)) %>%
  group_by(date) %>%
  summarise(N = sum(n))
p <- ggplot(ISR_con, aes(x = date, y = N)) +
  xlab(paste0("By ", dateBy)) +
  ylab("# of connections") +
  ggtitle("Number of ISR connections over time") +
  scale_x_date(limits = c(from, to)) +
  theme_IS(base_size = 18)
if (plotType == "line") {
  p <- p + geom_line()
} else {
  p <- p + geom_bar(stat = "identity")
}
ggplotly(p)
```

We had **`r round(mean(ISR_con$N), 2)` connections** per *`r dateBy`* on average from `r from` to `r to`.
<br>


```{r isr-user}
ISR_user <- ISR_inits %>% 
  select(date, X13) %>% 
  filter(X13 != "-") %>% 
  distinct() %>% 
  select(date) %>% 
  count(date) %>% 
  right_join(from_to, by = "date") %>% 
  mutate(n = replace(n, is.na(n), 0)) %>% 
  mutate(date = floor_date(date, dateBy)) %>%
  group_by(date) %>%
  summarise(N = sum(n))
p <- ggplot(ISR_user, aes(x = date, y = N)) +
  xlab(paste0("By ", dateBy)) +
  ylab("# of users") +
  ggtitle("Number of ISR users over time") +
  scale_x_date(limits = c(from, to)) +
  theme_IS(base_size = 18)
if (plotType == "line") {
  p <- p + geom_line()
} else {
  p <- p + geom_bar(stat = "identity")
}
ggplotly(p)
```

We had **`r round(mean(ISR_user$N), 2)` ISR users** per *`r dateBy`* on average from `r from` to `r to`.
<br>


## Downloads

```{r isr-bioc}
BIOC <- read.table("https://www.bioconductor.org/packages/stats/bioc/ImmuneSpaceR/ImmuneSpaceR_stats.tab", header = T)
BIOC <- BIOC %>% 
  filter(Month != "all") %>% 
  mutate(date = as.Date(paste(Year, Month, "01", sep = "-"), "%Y-%b-%d")) %>% 
  arrange(date) %>% 
  filter(date <= Sys.Date()) %>% 
  select(date, Nb_of_downloads)

p <- ggplot(BIOC, aes(x = date, y = Nb_of_downloads)) +
  xlab("By Month") +
  ylab("Number of downloads") +
  ggtitle("Number of downloads in Bioconductor over time") +
  scale_x_date(limits = c(from, to)) +
  theme_IS(base_size = 18)
if (plotType == "line") {
  p <- p + geom_line()
} else {
  p <- p + geom_bar(stat = "identity")
}
ggplotly(p)
```
<br>


```{r isr-github}
```


***

# RStudio
```{r rstudio}
log_rstudio <- logs_dt %>% 
  filter(grepl("GET /_rstudio/ HTTP/1.1", X6))


session_rstudio <- log_rstudio %>% 
  select(date) %>% 
  count(date) %>% 
  right_join(from_to, by = "date") %>% 
  mutate(n = replace(n, is.na(n), 0)) %>% 
  mutate(date = floor_date(date, dateBy)) %>%
  group_by(date) %>%
  summarise(N = sum(n))
p <- ggplot(session_rstudio, aes(x = date, y = N)) +
  xlab(paste0("By ", dateBy)) +
  ylab("# of sessions") +
  ggtitle("RStudio sessions over time") +
  scale_x_date(limits = c(from, to)) +
  theme_IS(base_size = 18)
if (plotType == "line") {
  p <- p + geom_line()
} else {
  p <- p + geom_bar(stat = "identity")
}
ggplotly(p)
```

We had **`r round(mean(session_rstudio$N), 2)` RStudio sessions** per *`r dateBy`* on average from `r from` to `r to`.
