# Evaluation of FlowCAP III Challenge 4 Cytotrol Results

<hr>

**Goal:** This report evaluates and compare the different algorithms that have participated to the FlowCAP IV challenge based on the Cytotrol data (T and B cell panels). 
The evaluation is using the population frequencies reported by the different group involved in the challenge. The table of statistics is available here.
Evaluation is done both in terms of variability across replicates/centers and bias compared to central manual analysis.

**Conclusion:** Based on the data shown below, we conclude that OpenCyto and flowDensity performs the best with relatively low variability and small bias compared to other algorithms. 
In addition, OpenCyto and flowDensity seemed to be complementary, in that one algorithm could outperform the other on some population statistics and vice-versa. For this reason, 
the two algorithms were combined as an OpenCyto pipeline, which was then used to analyze the Seracare data. A report peforming the automated gating for all panels can be found [here](/reports/HIPC/Lyoplate/runReport.view?reportId=module%3ALyoplateVisualization%2Freports%2Fschemas%2FAnalyses%2FAutomatedGatingOfSeracareData%28AllPanels%29.Rmd&redirectUrl=%2Fproject%2FHIPC%2FLyoplate%2Fbegin.view%3FpageId%3DAnalyses). The evaluation of the results 
   compared to manual analysis can be found [here](/study/HIPC/Lyoplate/dataset.view?Dataset.reportId=db%3A159&datasetId=5001).

<hr>

```{r knitr-opts, echo = FALSE, message = FALSE, cache = FALSE}
library(knitr)
opts_chunk$set(fig.align = "left", fig.width = 12, fig.height = 12)
```

```{r echo=FALSE,eval=TRUE}
path <- "/share/files/HIPC/Lyoplate/@files/scripts"
suppressPackageStartupMessages({
    library(data.table)
    library(car)
    library(lme4)
    library(plyr)
    library(reshape2)
    library(Rlabkey)
    co <- labkey.setCurlOptions(sslversion = 1)
    library(reshape)
    library(ggplot2)
    library(plotly)
})

source(file.path(path, "helperFunctions.r"))
fulldata <- labkey.selectRows(baseUrl = labkey.url.base, 
                              folderPath = "HIPC/Lyoplate", 
                              schemaName = "study", 
                              queryName = "Cytotrol FlowCAP III Data (T-cell and B-cell Panels)")
fulldata <- data.table(fulldata)
fulldata <- fulldata[, lp := logit(Proportion, adjust = 0)]
fulldata <- fulldata[, Method := factor(Method)]
```


## B-cell Panel

```{r bcell, eval = TRUE, warning = FALSE, message = FALSE}
mer <- .fitMer(fulldata[Panel == "Bcell"])
vc2 <- .extractVcomp(mer, sd = FALSE)
vc2 <- ddply(vc2, 
             .(Method, Population), 
             function(x) with(x, data.frame(variable, value = value / sum(value))))
vc <- .extractVcomp(mer, sd = TRUE)

vc2 <- .reorder.factor(vc2)
vc <- .reorder.factor(vc)
p <- ggplot(subset(vc, !is.nan(value))) +
  geom_bar(aes(x = Method, 
               y = value, 
               fill = variable), 
           stat = "identity", 
           position = "stack") +
  ggtitle("Variability per Method for B-Cell Panel") +
  facet_wrap(~Population, 
             scales = "free_y", labeller = label_both) +
  xlab("<br><br><br><br><br><br><br>Method") + 
  theme_bw() +
  theme(axis.title.x = element_text(size = 10, hjust = 0, vjust = 5),
        axis.title.y = element_text(size = 10),
        axis.text.x = element_text(angle = 45,vjust = 10),
        legend.position = "bottom") +
  scale_y_continuous("Standard Deviation") +
  scale_fill_brewer("Source of Variability", 
                    type = "qual")
p
ggplotly(p)



merb <- .fitMerBias(as.data.frame(fulldata[Panel %in% "Bcell"]))


cb <- .getCI(merb)
cytotrol_transci <- function(bar) {
  foo <- ddply(cast(melt(bar), 
                    Method + Population + variable ~ Sample), 
               .(Method, Population, variable), 
               function(x) with(x,
                                data.frame(Sample1 = `(Intercept)`, 
                                           Sample2 = `(Intercept)` + Sample2, Sample3 = `(Intercept)` + Sample3, Sample4 = `(Intercept)` + Sample4)))
  setnames(foo, "variable", "ci")
  foo <- melt(foo)
  foo$value <- arm::invlogit(foo$value)
  foo <- setnames(foo, "variable", "Sample")
  foo <- (cast(foo, Method + Population + Sample ~ ci))
  foo
}

cb <- cytotrol_transci(cb)
p <- ggplot(cb) +
  geom_point(aes(x = mean, 
                 y = Sample:Method, 
                 col = Sample, 
                 shape = Method) , 
             size = 3) +
  geom_segment(aes(x = lower, 
                   xend = upper, 
                   y = Sample:Method, 
                   yend = Sample:Method,
                   col = Sample), 
               lineend = "round", 
               lwd = 1) +
  theme_bw() +
  scale_x_continuous("Proportion") +
  scale_y_discrete() +
  facet_wrap(~Population, 
             ncol = 3, 
             scales = "free_x") +
  theme(axis.text.y = element_blank(), 
        axis.title.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        legend.position = "bottom",
        legend.direction = "horizontal") +
  scale_color_brewer(type = "qual") + 
  ggtitle("Bias for B-cell Panel (FlowCAP III)") + 
  xlim(c(0, 0.5))
p
ggplotly(p)
```
             
             
## T-cell Panel
             
```{r tcell, eval = TRUE, warning = FALSE, message = FALSE}

mer <- .fitMer(fulldata[Panel == "Tcell"])
vc2 <- .extractVcomp(mer, sd = FALSE)
vc2 <- ddply(vc2, 
             .(Method, Population), 
             function(x) with(x, data.frame(variable, value = value / sum(value))))
vc <- .extractVcomp(mer, sd = TRUE)

vc2 <- .reorder.factor(vc2)
vc <- .reorder.factor(vc)
p <- ggplot(subset(vc, !is.nan(value))) +
  geom_bar(aes(x = Method, 
               y = value, 
               fill = variable), 
           stat = "identity", 
           position = "stack") +
  ggtitle("Variability per Method for T-Cell Panel") +
  facet_wrap(~Population, 
#             scales = "free_y") + 
             nrow = 2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position = "bottom") +
  scale_y_continuous("Standard Deviation") +
  scale_fill_brewer("Source of<br> Variability", 
                    type = "qual")
p
ggplotly(p)

mert <- .fitMerBias(as.data.frame(fulldata[Panel %in% "Tcell"]))
ct <- .getCI(mert)
ct <- cytotrol_transci(ct)
ct$StdPop <- factor(ct$Population)
levels(ct$StdPop) <- c("CD4 Activated", 
                       "CD8 Activated", 
                       "CD4 Central Memory", 
                       "CD8 Central Memory", 
                       "CD4 Effector", 
                       "CD8 Effector", 
                       "CD4 Effector Memory", 
                       "CD8 Effector Memory", 
                       "CD4 Naive", 
                       "CD8 Naive")
ct$Population <- ct$StdPop

p <- ggplot(ct) +
  geom_point(aes(x = mean, 
                 y = Sample:Method, 
                 col = Sample, 
                 shape = Method)) +
  geom_segment(aes(x = lower, 
                   xend = upper,
                   y = Sample:Method, 
                   yend = Sample:Method, 
                   col = Sample), 
               lineend = "round") +
  theme_bw() +
  scale_x_continuous("Proportion") +
  scale_y_discrete() +
  facet_wrap(~Population, 
             scales = "free_x", 
             ncol = 3) +
  theme(axis.text.y = element_blank(), 
        axis.title.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        legend.position = "bottom", 
        legend.direction = "horizontal")+
  scale_color_brewer(type = "qual") + 
  ggtitle("Bias for T-cell Panel  (FlowCAP III)") + 
  xlim(c(0, 0.7))
p
ggplotly(p)
```

<script>
window.HTMLWidgets.staticRender();
</script>
