```{r init,echo=FALSE,eval=TRUE, message=FALSE}
library(knitr)
opts_chunk$set(dev = "CairoPNG",
               fig.align = "left",
               cache = FALSE,
               echo = FALSE,
               results = "hide",
               message = TRUE,
               warning = FALSE,
               error = TRUE)
```

### Import Manual Gates for Lyoplate Panel Data from FlowJo Workspaces (XML) and FCS Files

```{r load_libs,echo=FALSE,message=FALSE,warning=FALSE}
library(flowWorkspace)
library(data.table)
library(Rlabkey)
co <- labkey.setCurlOptions(sslversion = 1)
library(digest)
library(plyr)
library(tools)
```

This script parses and imports the manually gated FCS files from FlowJo workspaces for each panel and center.   
We use `flowWorkspace` to parse the XML workspaces and reproduce the manual gating using R's core flow infrastructure.
Click the `source` tab to see the code.
   
### Loading

```{r setup_paths_load_parse_and_write}
message("Setting up standard paths")
#' Set up some standard paths
scriptPath<-"/share/files/HIPC/Lyoplate/@files/scripts/"
path <- dirname(scriptPath)

# source some common functions
source(file.path(scriptPath, "wrapper-functions.r"))
nodes<-list()
gs_lists<-NULL
#' load panel-specific settings (uncomment the panel you want to analyze)
for(PANELSETTINGS in list.files(path=scriptPath,pattern="settings*")){
source(file.path(scriptPath,PANELSETTINGS))

#' Should we extract the spillover matrices for each center?
#' By default, no because the data are already compensated and transformed.
extract_spillover <- FALSE

#' whether to use ncdfFlowSet or flowSet
isNcdf <- TRUE

#' path that stores the manually gated FCS files and XML workspaces
manual_path <- path

#' flowJo workspace path
ws_path <- file.path(manual_path, "XML")

#' raw FCS file path
fcs_path <- file.path(manual_path, "FCS")

#' path that sores the phenoData annotations as excel file 
pd_path <- file.path(path, "pData")

message("Trying to  load saved workspaces")
# Test if the data has already been parsed by trying to load the gatingset list
if(inherits(try(suppressMessages(gs_list<-load_gslist(gslist_path_manual))),"try-error")){
  message("Can't find manually gated gslist. Re-importing:");
  
centers <- c('BIIR','CIMR','Miami','NHLBI','Stanford','UCLA','Yale')
  
#' If specified, extracts spillover matrices for each center as a list
if (extract_spillover) {
  message("Extracting spillover matrices...")
  spillover_matrices <- sapply(centers, function(center) {
    spillover_file <- file.path(ws_path, center, paste(center, "Matrices"), panel)
    read.table(spillover_file, skip = 2, nrows = 8, header = TRUE, check.names = FALSE, sep = "\t")
  }, simplify = FALSE)
}


#' We parse each center's workspace and then extract its flowSet.
message("Parsing Lyoplate workspaces...")
gs_list <- sapply(centers, function(center) {
  message("Center: ", center)
  this_ws_path <- file.path(ws_path, center, paste0("CA c3 v2 ", center, ".xml"))
  ws <- openWorkspace(this_ws_path)
  groupNames <- levels(getSampleGroups(ws)[,"groupName"])
  #NOTE: here we use the fuzzy match due to the short name used in auto gating
  #ideally, we want to do the strict full string match avoid picking up the wrong panel
  #in case the panel name are similar. But in this particular panel set, it is safe to do so.    
  groupID <- grep(panel, groupNames, ignore.case = T)
  #NOTE: in theory, we shouldn't need to do this if center names were consistent.
    
  this_fcs_path <- file.path(fcs_path, center)
  this_fcs_path <- sub("BIIR", "Baylor", this_fcs_path)
  gs <- parseWorkspace(ws, name = groupNames[groupID], path = this_fcs_path, isNcdf = isNcdf, extend_val = -10)
  gs
}, simplify = FALSE)


# Swaps the channels and markers for the current 'flowSet' object. This ensures
# that we can 'rbind2' the 'GatingSetList' below because the stain names do not
# match otherwise.
message ("Swapping flowSet channels and markers")
gs_list <- sapply(centers, function(center) {
  message("Center: ", center)

  gs <- gs_list[[center]]
  fs <- getData(gs)

  for(sn in sampleNames(fs)){
      fr <- fs@frames[[sn]]
      fs@frames[[sn]] <- preprocess_flowframe(fr, use.exprs = !isNcdf)
  }

  newColNames <- colnames(fs@frames[[sn]])
  colnames(fs) <- newColNames #assuming the order of colnames between fr and fs were consistent
  flowData(gs) <- fs
  
  fr <- fs@frames[[sn]]
  pd <- pData(parameters(fr))
  pd <- pd[!is.na(pd$desc), 2:1]
  colnames(pd) <- c("old", "new")

  gs <- updateGateParameter(gs, pd)
  
  #subset on common columns
  fs <- fs[, markers_of_interest]
  flowData(gs) <- fs
  gs
}, simplify = FALSE)

#' hide singlet gate since some centers don't have it
invisible(lapply(centers[!centers%in%centers_no_singlets], function(center){
          setNode(gs_list[[center]], "singlets", FALSE)
        }))

#' correct typos for some centers
invisible(lapply(names(popName_alias), function(thisCenter){
          message(thisCenter)
          gs <- gs_list[[thisCenter]]
          typo <- popName_alias[[thisCenter]]
          lapply(names(typo), function(thisTypo){
                correct <- typo[[thisTypo]]
                message(thisTypo, " --> ", correct)
                setNode(gs, thisTypo, correct)      
              })
          
        }))
  
#' Construct a gating set list  
gs_list <- GatingSetList(gs_list)

# Updates pData to include Center, Sample, and Replicate
# To do this, we open the csv files provided for each center and extract the
# relevant data.
message("Extracting pData from csv config files...")
center_pData <- ldply(centers, function(center) {
  message("Center: ", center)
  
  #NOTE: in theory, we shouldn't need to do this if center names were consistent.
  center <- sub("BIIR", "Baylor", center)
  # The filename of the manually edited Excel file
  csv <- dir(path = pd_path, pattern = center, full.names = TRUE)
  
  exp_samples <- read.csv(file = csv, stringsAsFactors = FALSE)
  colnames(exp_samples) <- c("name", "Institution", "Panel", "Replicate",
                             "Sample")
  exp_samples <- subset(exp_samples, select = -c(Institution, Panel))
  exp_samples$Replicate <- as.character(as.numeric(exp_samples$Replicate))
  exp_samples$Sample <- as.character(as.numeric(exp_samples$Sample))
  exp_samples$Center <- center
  exp_samples
})

center_pData <- subset(center_pData, select = c(name, Center, Sample, Replicate))

message("Updating flowSet's pData...")
center_pData <- merge(pData(gs_list), center_pData, sort = FALSE)
rownames(center_pData) <- rownames(pData(gs_list))
pData(gs_list) <- center_pData


#' Archives the results
save_gslist(gs_list, path = gslist_path_manual)
}
  
#'Rename to commmon names and save those too:
suppressMessages(gslist <- load_gslist(gslist_path_manual))
rename_pops(gslist, subpopulations_manual, subpopulations_common)
rename_pops(gslist, "not dead", "Live")
save_gslist(gslist, path = sub("gated_data", "gated_data/pop_renamed", gslist_path_manual), cdf = "link", overwrite = T)


#' extract stats from manual gates
pop_stats <- extractStats(gslist, subpopulations_common)

#Check the md5sum of the current data and the existing data. Re-generate if it's stale.

md5current <- md5sum(list.files(path=dirname(stats_path_manual),pattern=basename(stats_path_manual),full=TRUE))
temp <- file.path(tempdir(),basename(stats_path_manual))

# Make the stats amenable to import into the database
pop_stats<-cbind(pop_stats,Panel=panel,Method="manual")
pop_stats<-data.table(pop_stats)
pop_stats=pop_stats[,c("ParticipantID","SequenceNum"):=  list(paste0(Filename,Panel,Center,Method,Sample,Population),1)]
pop_stats[,ParticipantID:=digest(ParticipantID),ParticipantID]
#rename the population levels
pop_stats$Panel=mapvalues(pop_stats$Panel,from=c("bcell","tcell","DC","treg"),to=c("Bcell","Tcell","DC","Treg"))
write.csv(pop_stats, file = temp, row.names = FALSE)
md5new <- md5sum(temp)
if(TRUE){#md5current!=md5new){
  message("Writing population statistics")
write.csv(pop_stats, file = stats_path_manual, row.names = FALSE)
   #query
   todelete<- labkey.selectRows(baseUrl=labkey.url.base,folderPath="HIPC/Lyoplate",schemaName="study",queryName="cell_population_statistics",colNameOpt="rname",showHidden=TRUE)
   todelete<-subset(todelete,method=="manual"&panel==unique(pop_stats$Panel))[,"lsid",drop=FALSE]
   #Delete and update
   if(nrow(todelete)>0){
#       labkey.deleteRows(baseUrl=labkey.url.base,folderPath="HIPC/Lyoplate",schemaName="study",queryName="cell_population_statistics",toDelete=todelete)
    }
   #Insert into labkey database
#   labkey.importRows(baseUrl=labkey.url.base,folderPath="HIPC/Lyoplate",schemaName="study",queryName="cell_population_statistics",toImport=as.data.frame(pop_stats,stringsAsFactors=FALSE))

}else{
  message("Population statistics already written")
  nodes<-c(nodes,list(getNodes(gslist[[1]])))
  gs_lists<-c(gs_lists,gslist)
}
message("Finished parsing manual gates for the ",panel," panel")
nodes<-c(nodes,list(getNodes(gslist[[1]])))
gs_lists<-c(gs_lists,gslist)
}
   
```
   
### Results
   
### `r gsub("settings-(.+)\\.r","\\1",list.files(path=scriptPath,pattern="settings*")[1])` Panel
   
```{r report1,results='markdown'}
PANELSETTINGS<-1
    message(paste0("The following cell populations are defined in the manual gates for the ",gsub("settings-(.+)\\.r","\\1",list.files(path=scriptPath,pattern="settings*")[PANELSETTINGS])," panel"))
    print(nodes[[PANELSETTINGS]])
   message("Gating Tree")
   plot(gs_lists[[PANELSETTINGS]][[1]])
   message("Example Gating Layout")
   plotGate(gs_lists[[PANELSETTINGS]][[1]])

```

### `r gsub("settings-(.+)\\.r","\\1",list.files(path=scriptPath,pattern="settings*")[2])` Panel

```{r report2,results='markdown'}
PANELSETTINGS<-2
    message(paste0("The following cell populations are defined in the manual gates for the ",gsub("settings-(.+)\\.r","\\1",list.files(path=scriptPath,pattern="settings*")[PANELSETTINGS])," panel"))
    print(nodes[[PANELSETTINGS]])
   message("Gating Tree")
   plot(gs_lists[[PANELSETTINGS]][[1]])
   message("Example Gating Layout")
   plotGate(gs_lists[[PANELSETTINGS]][[1]])

```

### `r gsub("settings-(.+)\\.r","\\1",list.files(path=scriptPath,pattern="settings*")[3])` Panel


```{r report3,results='markdown'}
PANELSETTINGS<-3

message(paste0("The following cell populations are defined in the manual gates for the ",gsub("settings-(.+)\\.r","\\1",list.files(path=scriptPath,pattern="settings*")[PANELSETTINGS])," panel"))
    print(nodes[[PANELSETTINGS]])
   message("Gating Tree")
   plot(gs_lists[[PANELSETTINGS]][[1]])
   message("Example Gating Layout")
   plotGate(gs_lists[[PANELSETTINGS]][[1]])

```

### `r gsub("settings-(.+)\\.r","\\1",list.files(path=scriptPath,pattern="settings*")[4])` Panel


```{r report4,results='markdown'}
PANELSETTINGS<-4

message(paste0("The following cell populations are defined in the manual gates for the ",gsub("settings-(.+)\\.r","\\1",list.files(path=scriptPath,pattern="settings*")[PANELSETTINGS])," panel"))
    print(nodes[[PANELSETTINGS]])
   message("Gating Tree")
   plot(gs_lists[[PANELSETTINGS]][[1]])
   message("Example Gating Layout")
   plotGate(gs_lists[[PANELSETTINGS]][[1]])

```

